---
title: "Yield Data Acquisition"
output: html_notebook
---

**Installing packages**

```{r}

library(tidyUSDA)
library(tidyverse)

```

We will acquire our yield data from the USDA Quick Stats database using a API key 
and the R package tidyUSDA. Please follow this link to get your own free API key 
https://quickstats.nass.usda.gov/api/ 

Save the API key to a variable known as key

```{r}

key <- "your API key"


years <- c("1976","1977","1978","1979","1980","1981","1982","1983","1984",
           "1985","1986","1987","1988","1989","1990","1991","1992","1993",
           "1994","1995","1996","1997","1998","1999","2000","2001","2002",
           "2003","2004","2005","2006","2007","2008","2009","2010","2011",
           "2012","2013","2014","2015","2016","2017","2018","2019","2020",
           "2021","2022")


Sunflower_Yield <- getQuickstat(sector='CROPS',
                        group = "FIELD CROPS",
                      commodity = "SUNFLOWER",
                      category = "YIELD",
                      domain = "TOTAL",
                      key = key,
                      program = 'SURVEY',
              data_item = "SUNFLOWER, OIL TYPE - YIELD, MEASURED IN LB / ACRE",
                geographic_level = 'COUNTY',
                   year = years)

```


Viewing what the data looks like after the download (Looking at the counties and
viewing how many times they are represented)

Then removing some of the states and counties outside of our study. We removed
California because of too few observations and removed the counties without
specific names

```{r}

Sunflower_Yield %>% dplyr::group_by(county_name) %>% summarise(n()) %>% View()

 
Sunflower_Yield <- Sunflower_Yield %>% 
                   dplyr::filter(county_name != "OTHER (COMBINED) COUNTIES") %>%
                   dplyr::filter(county_name != "OTHER COUNTIES") %>%
                   dplyr::filter(state_name != "CALIFORNIA") %>%
                   dplyr::select(year,state_name,county_name,Value) %>%
                   dplyr::rename(Yield = Value)





```


We are going to be doing the same thing with the Acres harvested and Acres Planted
data i.e., acquire the data, then remove California and counties without specific
names. 

```{r}


########### AREA HARVESTED DATA ##### 

Area_Harvested <- getQuickstat(sector='CROPS',
                                group = "FIELD CROPS",
                                commodity = "SUNFLOWER",
                                category = "AREA HARVESTED",
                                domain = "TOTAL",
                                key = key,
                                program = 'SURVEY',
                                data_item = "SUNFLOWER, OIL TYPE - ACRES HARVESTED",
                                geographic_level = 'COUNTY',
                                year = years)


### removing some of the states and counties outside of our study 

Area_Harvested <- Area_Harvested %>% 
                  dplyr::filter(county_name != "OTHER (COMBINED) COUNTIES") %>%
                  dplyr::filter(county_name != "OTHER COUNTIES") %>%
                  dplyr::filter(state_name != "CALIFORNIA") %>%
                  dplyr::select(year,state_name,county_name,Value) %>%
                  dplyr::rename(Acres_Harvested = Value)



###### AREA PLANTED DATA #### 

Area_Planted <- getQuickstat(sector='CROPS',
                               group = "FIELD CROPS",
                               commodity = "SUNFLOWER",
                               category = "AREA PLANTED",
                               domain = "TOTAL",
                               key = key,
                               program = 'SURVEY',
                               data_item = "SUNFLOWER, OIL TYPE - ACRES PLANTED",
                               geographic_level = 'COUNTY',
                               year = years)


### removing some of the states and counties outside of our study 

Area_Planted <- Area_Planted %>% 
                  dplyr::filter(county_name != "OTHER (COMBINED) COUNTIES") %>%
                  dplyr::filter(county_name != "OTHER COUNTIES") %>%
                  dplyr::filter(state_name != "CALIFORNIA") %>%
                  dplyr::select(year,state_name,county_name,Value) %>%
                  dplyr::rename(Acres_Planted = Value)


```

**Joining the Yield data, Acres Planted and Acres Harvested data**

```{r}

##### Joining these three dataframes into one ### 

Sunflower_Yield <- list(Area_Planted,Area_Harvested,Sunflower_Yield) %>% 
                   reduce(inner_join)


```


**States and counties in in my study**

```{r}

######## Viewing the oldest and latest values for each counties ###
### Identifying which counties have yield values on the either side of the year 2005

## Study counties ###
Counties <- Sunflower_Yield %>% 
                  group_by(state_name,county_name) %>% 
                  summarise(Latest = max(year),
                            Oldest = min(year)) %>% 
                          arrange(desc(Latest)) %>% 
                          filter(Latest >= 2005,
                                 Oldest <= 2004)

##### Filtering the Sunflower Yield dataframe based on these counties 

Sunflower_Yield <- Sunflower_Yield %>% 
                            inner_join(Counties)



####  removing rows containing Yield values = 0 ##

Sunflower_Yield <- Sunflower_Yield %>% 
                        filter(Yield > 0)

```


**Exporting datasets**

```{r}

########################
## EXPORTING THE FILES IN A FOLDER CALLED Raw_Datasets and in a subfolders called Yield and Regions respectively##
#####################


### write out the yield and the region file ### 

## export to subfolder Yield

write.csv(Sunflower_Yield,
          "~/Yield_forecast_Sunflowers/Yield_forecast_Sunflowers/Raw_Datasets/Yield/Sunflower_Yield.csv",
          row.names = FALSE)

## export to subfolder Regions

write.csv(Counties,
          "~/Yield_forecast_Sunflowers/Yield_forecast_Sunflowers/Raw_Datasets/Regions/Study_Regions.csv",
          row.names = F)   

```



























