---
title: "Future Climate Data acquisition"
output: html_notebook
---

Future Climate data used in this study was downloaded from WorldClim database
(https://www.worldclim.org/data/cmip6/cmip6_clim2.5m.html).Data from ACCESS-ESM1-5
climate model at 2.5 minutes spatial resolution was used.

The datasets used are named as tn (for minumum temperature),
tx (for maximum temperature) and pr (for precipitation).


Alternatively, here is some code to acquire the data using the R package geodata

**Loading the libraries**

```{r}

library(here)
library(geodata)


```


Create a subfolder within the Raw_Datasets folder 


```{r}

######### 2021-2040 ####

######## Tmin

path = "path to data"

cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2021-2040",
            var = "tmin", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2021-2040",
            var = "tmin", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2021-2040",
            var = "tmin", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2021-2040",
            var = "tmin", res = 2.5, 
            path = path) 



####### Tmax 

cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2021-2040",
            var = "tmax", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2021-2040",
            var = "tmax", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2021-2040",
            var = "tmax", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2021-2040",
            var = "tmax", res = 2.5, 
            path = path) 



####### Precipitation 

cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2021-2040",
            var = "prec", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2021-2040",
            var = "prec", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2021-2040",
            var = "prec", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2021-2040",
            var = "prec", res = 2.5, 
            path = path) 





###################### 2041-2060 ############ 

######## Tmin

path = "path to data"

cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2041-2060",
            var = "tmin", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2041-2060",
            var = "tmin", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2041-2060",
            var = "tmin", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2041-2060",
            var = "tmin", res = 2.5, 
            path = path) 

######## Tmax


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2041-2060",
            var = "tmax", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2041-2060",
            var = "tmax", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2041-2060",
            var = "tmax", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2041-2060",
            var = "tmax", res = 2.5, 
            path = path) 


##### Precipitation  


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2041-2060",
            var = "prec", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2041-2060",
            var = "prec", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2041-2060",
            var = "prec", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2041-2060",
            var = "prec", res = 2.5, 
            path = path) 



############## 2061-2080 #############

######## Tmin

path = "path to data"

cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2061-2080",
            var = "tmin", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2061-2080",
            var = "tmin", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2061-2080",
            var = "tmin", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2061-2080",
            var = "tmin", res = 2.5, 
            path = path) 


########## Tmax 


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2061-2080",
            var = "tmax", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2061-2080",
            var = "tmax", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2061-2080",
            var = "tmax", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2061-2080",
            var = "tmax", res = 2.5, 
            path = path) 


#### Precipitation 


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '126', time = "2061-2080",
            var = "prec", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '245', time = "2061-2080",
            var = "prec", res = 2.5, 
            path = path)



cmip6_world(model = 'ACCESS-ESM1-5',ssp = '370', time = "2061-2080",
            var = "prec", res = 2.5, 
            path = path)


cmip6_world(model = 'ACCESS-ESM1-5',ssp = '585', time = "2061-2080",
            var = "prec", res = 2.5, 
            path = path) 


```